Verifying Spatial Properties of Array Computations

### 2017-04-15 oopsla17/study.tex
improved table 2

\item Many array computations are \emph{stencil computations},
with a static access pattern as described in Hypothesis 2, writing
to an array at an index based on a (possibly constant offset from induction
variables, \eg{}, the classic five-point stencil;

@@

\noindent
We revist each hypothesis and show the related data and conclusions.
%
\begin{enumerate}
\item Loops over arrays mainly read from arrays with a static pattern
based on constant offsets from (base or dervied) induction variables;
\end{enumerate}
%
We identified $108,773$ instances of assignments within loops in
which an array subscript flows to the right-hand side. We refer
to each one of these as an \emph{array computation}.
We performed the classification described above
and grouped the data by the right-hand side classification only:
%%
\begin{center}
\begin{tabular}{lS[table-format=4]@{\extracolsep{5pt}}S[table-format=2.2]c}
\textbf{RHS classification} & \textbf{Number} & \% &
\textit{grouping 1} \\ \hline
Affine                          & 34        & 0.03\%
& \multirow{3}{*}{81.43\%} \\ \cline{1-3}
Neighbourhood                   & 65334     & 60.06 \%  \\ \cline{1-3}
Neighbourhood + constants       & 23209     & 21.34\%  \\ \hline
Other                           & 20196     & 18.57 \%  \\ \hline \hline
Total                           & 108773    &  \\
\end{tabular}
\end{center}
%
\noindent
Since, affine and neighbour patterns are static, we
group them (\textit{grouping 1}) to see that $81.43\%$ of the array
computations (88,577) have a static pattern that are either constants or
constant offsets of base/derived induction variables, confirming
\emph{hypothesis 1}. Note that affine (derived induction variables)
are rare. Overwhelmingly, the main category is neighbourhood
offsets (possibly mixed with constants in some
dimensions). The ``affine + constants'' class is
not represented at all.
%
\begin{enumerate}[resume]
\item Most loop-array computations of the previous form read
from arrays with a \emph{contiguous} pattern, \eg{}:
%
\begin{minted}{fortran}
do i = 1, n
  b(n) = (a(i) + a(i+1)) / 2.0
\end{minted}
%
\item Most loop-array computations of the previous form read
from arrays with a pattern that includes the immediate
neighbours (offsets of 1 from the induction variables);
\end{enumerate}
%
We grouped the data based on subclassifications of \emph{grouping 1}
in terms of contiguity and positioning.
%Focussing on grouping 1.
\begin{center}
\begin{tabular}{lS[table-format=4]S[table-format=2.2]S[table-format=2.2]lS[table-format=4]S[table-format=2.2]S[table-format=2.2]}
\textbf{RHS classification} & \textbf{Number} & \% {(of whole)}
                                              & \% {(of group)}
&\textbf{Position} & \textbf{Number} & \% {(whole)} & \% {(group)} \\ \hline
Affine, contiguous
& 31 & 0.03\% & 0.03\% \\ \hline
%\multirow{3}{*}{Affine, contiguous}
% & %\multirow{3}{*}{31}
% & %\multirow{3}{*}{0.03\%}
% & \textsf{away} & 4 & 0.00\% \\
%& 31 & 0.03\% & \textsf{origin} & 15 & 0.01\% \\
%&&& \textsf{straddle} & 12 & 0.01\% \\ \hline
Affine, non-contiguous
 & 3
 & 0.00\% & 0.00\% \\ \hline
\multirow{3}{*}{Neighbour, contiguous}
 & %\multirow{3}{*}{83689}
 & %\multirow{3}{*}{83.08\%}
& & \textsf{away} & 152 & 0.14\% & 0.17 \% \\
& 83689 & 76.94\% & 94.48\% & \textsf{origin} & 81309 & 74.75\% & 91.79\% \\
&&&& \textsf{straddle} & 2228 & 2.05\% & 2.52\% \\ \hline
Neighbour, non-contiguous
 & 4854
 & 4.46\% & 5.48\%  \\ \hline \hline
\textbf{Total}  & 88577    &  \\
\end{tabular}
\end{center}
%%
Thus, $76.94\%$ of array computations have only contiguous
neighbour indices (which may include constants in some indices)
in a contiguous pattern on the right hand side, with $76.80\%$
either at the origin ($74.75\%$) or offset by a distance of 1 from the origin
($2.05\%$). This confirms both hypotheses 2 and 3. Only 152 instances
were found of index patterns not immediate next to or over the origin.

>>>>

\item Many array computations are \emph{stencil computations}: an
  assignment whose right-hand side comprises array subscripts with
  neighbourhood indices and whose left-hand side is an
  neighbourhood-indexed array subscript (\eg{}, the classic five-point
  stencil);

@@

\noindent
We revist each hypothesis and show the related data and conclusions.
%
\begin{enumerate}
\item Loops over arrays mainly read from arrays with a static pattern
based on constant offsets from (base or dervied) induction variables;
\end{enumerate}
%
We identified $108,773$ instances of assignments within loops in
which an array subscript flows to the right-hand side. We refer
to each one of these as an \emph{array computation}.
We performed the classification described above
and grouped the data by the right-hand side classification only:
%%
\begin{center}
\begin{tabular}{lS[table-format=4]@{\extracolsep{5pt}}S[table-format=2.2]c}
\textbf{RHS classification} & \textbf{Number} & \% &
\textit{grouping 1} \\ \hline
Affine                          & 34        & 0.03\%
& \multirow{3}{*}{81.43\%} \\ \cline{1-3}
Neighbourhood                   & 65334     & 60.06 \%  \\ \cline{1-3}
Neighbourhood + constants       & 23209     & 21.34\%  \\ \hline
Other                           & 20196     & 18.57 \%  \\ \hline \hline
Total                           & 108773    &  \\
\end{tabular}
\end{center}
%
\noindent
Since, affine and neighbour patterns are static, we
group them (\textit{grouping 1}) to see that $81.43\%$ of the array
computations (88,577) have a static pattern that are either constants or
constant offsets of base/derived induction variables, confirming
\emph{hypothesis 1}. Note that affine (derived induction variables)
are rare. Overwhelmingly, the main category is neighbourhood
offsets (possibly mixed with constants in some
dimensions). The ``affine + constants'' class is
not represented at all.
%
\begin{enumerate}[resume]
\item Most loop-array computations of the previous form read
from arrays with a \emph{contiguous} pattern, \eg{}:
%
\begin{minted}{fortran}
do i = 1, n
  b(n) = (a(i) + a(i+1)) / 2.0
\end{minted}
%
\item Most loop-array computations of the previous form read
from arrays with a pattern that includes the immediate
neighbours (offsets of 1 from the induction variables);
\end{enumerate}
%
We grouped the data based on subclassifications of \emph{grouping 1}
in terms of contiguity and positioning.
%Focussing on grouping 1.
\begin{center}
\begin{tabular}{lS[table-format=4]S[table-format=2.2]lS[table-format=4]S[table-format=2.2]}
\textbf{RHS classification} & \textbf{Number} & \% {(of whole)}
&\textbf{Position} & \textbf{Number} & \% {(of whole)} \\ \hline
Affine, contiguous
& 31 & 0.03\% \\ \hline
%\multirow{3}{*}{Affine, contiguous}
% & %\multirow{3}{*}{31}
% & %\multirow{3}{*}{0.03\%}
% & \textsf{away} & 4 & 0.00\% \\
%& 31 & 0.03\% & \textsf{origin} & 15 & 0.01\% \\
%&&& \textsf{straddle} & 12 & 0.01\% \\ \hline
Affine, non-contiguous
 & 3
 & 0.00\% & \\ \hline
\multirow{3}{*}{Neighbour, contiguous}
 & %\multirow{3}{*}{83689}
 & %\multirow{3}{*}{83.08\%}
 & \textsf{away} & 152 & 0.14\% \\
& 83689 & 76.94\% & \textsf{origin} & 81309 & 74.75\% \\
&&& \textsf{straddle} & 2228 & 2.05\% \\ \hline
Neighbour, non-contiguous
 & 4854
 & 4.46\% \\ \hline
Other & 20196 & 18.57\%  \\
 \hline
\textbf{Total}  & 88577    &  \\
\end{tabular}
\end{center}
%%
Thus, $76.94\%$ of array computations have only contiguous neighbour
indices (which may include constants in some indices) in a contiguous
pattern on the right hand side, with $76.80\%$ either at the origin
($74.75\%$) or offset by a distance of 1 from the origin
($2.05\%$). This confirms both hypotheses 2 and 3. Only 152 instances
of index patterns not immediately next to or over the origin were
found.  There is a non trivial amount of non-contiguous arrays access,
but it is a much smaller proportion of the whole.

\begin{enumerate}[resume]
\item Many array computations are \emph{stencil computations}:
an assignment whose
 right-hand side comprises array subscripts with neighbourhood indices
and whose left-hand side is an neighbourhood-indexed array subscript
(\eg{}, the classic five-point stencil);
\end{enumerate}
